On The Difficulty of Computing Non-Commutative Signature Gröbner Bases

### 2022-02-07 article_issac.tex
Def de l'alphabet pour la prop sur les langages réguliers.
section context free

Looking at principal syzygies and the algorithm for computing new ones, we wonder if an automaton can recognize leading monomials of syzygies in the case where the Gröbner basis is finite. Let us show that this is not the case using the non regular language\linebreak $\{a^nb^n \;|\; n \in \N\}$. Let us consider the alphabet 
$$\{a, b, c, d, e, s, t, x, y, z, \infty\}$$
and polynomials
\begin{align*}
g_1 & := ssa - tcx \\
g_2 & := xa - ax \\
g_3 & := xd - dx \\
g_4 & := ya - ay \\
g_5 & := yd - dy \\
g_6 & := xbb - dyb \\
g_7 & := cya - ccx \\
g_8 & := xbe - dze \\
g_9 & := dz - zb \\
g_{10} & := cz - za \\
g_{11} & := ss-tz
\end{align*}
together with the order $m_1 < m_2$ if:
\begin{itemize}
\item the degree of $m_1$ is strictly smaller than the degree of $m_2$,
\item $m_1$ and $m_2$ have the same degree and there are less $\infty$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same degree, the same number of $\infty$ symbols and there are less $s$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same degree, the same numbers of $\infty$, and the same number $s$ symbols, but
$m_1 <_{lex} m_2$ where\\ $a < b < c < d < x < y < z < e < t < s < \infty$.
\end{itemize}
Such defined, this order is fair.

@@

\begin{align*}
g_1 & := ssa - t\alpha x \\ 
g_2 & := xa - ax \\ 
g_{11} & := x\beta-\beta x\\ 
g_{12} & := xb-\beta y \\ 
%g_3 & := xbb - \beta yb\\ 
% Inutile, redondante avec g_12
g_4 & := yb-by\\ 
g_{13} & := y\gamma-\gamma y \\
g_5 & := ycc-\gamma zc \\ 
g_6 & :=  \gamma z-z\gamma\\ 
g_7 & := bz-zb \\
g_8 & := \beta z-z\beta \\
g_9 & := az-za \\ 
g_{10} & := \alpha za-\alpha\alpha x  \\ 
g_{19} & := yce - \gamma oe \\
g_{14} & := \gamma o-oc \\
%g_{15} & := oc-\gamma o \\ 
% Inutile, redondante avec g_14
g_{16} & := \beta o-ob \\ 
g_{17} & := \alpha o-oa \\ 
g_{18} & := to-ss
\end{align*}

>>>>

Looking at principal syzygies and the algorithm for computing new ones, we wonder if an automaton can recognize leading monomials of syzygies in the case where the Gröbner basis is finite. Let us show that this is not the case using the non regular language\linebreak $\{a^nb^n \;|\; n \in \N\}$. Let us consider the alphabet 
$$X:=\{a, b, c, d, e, s, t, x, y, z, \infty\}$$
and polynomials
\begin{align*}
g_1 & := ssa - tcx \\
g_2 & := xa - ax \\
g_3 & := xd - dx \\
g_4 & := ya - ay \\
g_5 & := yd - dy \\
g_6 & := xbb - dyb \\
g_7 & := cya - ccx \\
g_8 & := xbe - dze \\
g_9 & := dz - zb \\
g_{10} & := cz - za \\
g_{11} & := ss-tz
\end{align*}
together with the order $m_1 < m_2$ if:
\begin{itemize}
\item the degree of $m_1$ is strictly smaller than the degree of $m_2$,
\item $m_1$ and $m_2$ have the same degree and there are less $\infty$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same degree, the same number of $\infty$ symbols and there are less $s$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same degree, the same numbers of $\infty$, and the same number $s$ symbols, but
$m_1 <_{lex} m_2$ where\\ $a < b < c < d < x < y < z < e < t < s < \infty$.
\end{itemize}
Such defined, this order is fair.

@@

We extend the result of the previous section by showing that even a pushdown automaton cannot recognize leading monomials of syzygies for a finite Gröbner basis. For that, we use the non algebraic language $\{a^nb^nc^n \;|\; n \in \N\}$. Let us consider the new alphabet 
$$X:=\{a, b, c, \alpha ,\beta , \gamma, e, s, t, x, y, z, o, \infty\}$$
and polynomials \fixmecc{pourquoi cette numérotation?}

\begin{align*}
g_1 & := ssa - t\alpha x \\ 
g_2 & := xa - ax \\ 
g_{11} & := x\beta-\beta x\\ 
g_{12} & := xb-\beta y \\ 
%g_3 & := xbb - \beta yb\\ 
% Inutile, redondante avec g_12
g_4 & := yb-by\\ 
g_{13} & := y\gamma-\gamma y \\
g_5 & := ycc-\gamma zc \\ 
g_6 & :=  \gamma z-z\gamma\\ 
g_7 & := bz-zb \\
g_8 & := \beta z-z\beta \\
g_9 & := az-za \\ 
g_{10} & := \alpha za-\alpha\alpha x  \\ 
g_{19} & := yce - \gamma oe \\
g_{14} & := \gamma o-oc \\
%g_{15} & := oc-\gamma o \\ 
% Inutile, redondante avec g_14
g_{16} & := \beta o-ob \\ 
g_{17} & := \alpha o-oa \\ 
g_{18} & := to-ss
\end{align*}
As in the previous section, we consider the fair order that first compares degrees, then the number of $\infty$ symbols, then the number of $s$ symbols and finally the lex order for\\
$a<b<c<\alpha<\beta<\gamma<x<y<z<o<e<t<s<\infty$. 
\fixmecc{vérifier}
For $n=2$, we get the syzygy with leadong monomial $\epsilon_{18}ssaabbcce$:

\begin{align*}
ssaabbcce & =_{1} t\alpha xabbcce =_{2} t\alpha axbbcce =_{12} t\alpha a\beta ybcce \\
& =_{4} t\alpha a\beta bycce =_{5} t\alpha a\beta b\gamma zce =_{6} t\alpha a\beta bz\gamma ce \\
& =_{7} t\alpha a\beta zb\gamma ce =_{8} t\alpha az\beta b\gamma ce =_{9} t\alpha za\beta b\gamma ce \\
& =_{10} t\alpha\alpha x\beta b\gamma ce =_{11} t\alpha\alpha\beta xb\gamma ce =_{12} t\alpha\alpha\beta\beta y\gamma ce \\
& =_{13} t\alpha\alpha\beta\beta\gamma yce =_{19} t\alpha\alpha\beta\beta\gamma\gamma oe =_{14}
t\alpha\alpha\beta\beta\gamma oce\\
& =_{14} t\alpha\alpha\beta\beta occe =_{16} t\alpha\alpha\beta obcce
=_{16} t\alpha\alpha obbcce\\ 
& =_{17}t\alpha oabbcce =_{17} toaabbcce =_{18}ssaabbcce
\end{align*}
\fixmecc{vérifier}

\begin{proposition}
  Leading monomials of syzygies of $F=\{g_1,\dots,g_{18}\}$ do not form an context free language.
  \label{prop:exp_pas_regulier}
\end{proposition}
\begin{proof}
  For every $n \geq 1$, $g_{11}a^nb^ne$ is the leading monomial of the syzygy:
